Marlowe: financial contracts on blockchain

### 2018-05-17 tex/paper.tex
implementation section

We make the following contributions in this paper.
\begin{itemize}
\item
Designing a DSL for financial contracts on blockchains: Marlowe.
\item
Defining an executable, small-step semantics of Marlowe in Haskell.
\item 
Embedding Marlowe as a Haskell \haskellinline{data} type to extend its expressibility.
\item
Developing the Meadow tool that allows users to interact with and simulate the operation of embedded Marlowe contracts. 
\end{itemize}

TO ADD
\begin{itemize}
\item 
Model vs implementation: add a section on the letter e.g. about distributed computation.
\item
In particular we're agnostic about UTxO and accounts, but want to be implementable on both.
\item
focus on some of the design decisions, but focus on the important aspects only and cut the stuff that's  more speculative/internal
\item  
\ldots
\end{itemize}

@@

we begin in Section \ref{section:design} with a discussion of the design rationale 
for Marlowe, including an overview of the various approaches to scripting in blockchain as well as the different designs 
of the blockchains themselves. We conclude Section \ref{section:design} with a set of decisions about the blockchain 
environment for Marlowe.

@@

\section{Design rationale}
\label{section:design}

\medskip
\noindent
In this section, we look at existing work on smart contracts and blockchains in general, before exploring the rationale 
for building the Marlowe language.

\subsection{Blockchain platforms: Cardano and others}

There is not a single model of how a blockchain operates, and in this section we look at some of the tradeoffs possible 
in this space. This allows us to give a context to some of the design decisions taken in designing Marlowe.

The simplest blockchain is Bitcoin. Transactions on Bitcoin are made by spending the (as yet) unspent outputs of 
previous transactions (`unspent transaction outputs' or UTxOs). The chain need not maintain any state, such as the 
current value owned by any particular participant: such \emph{account} information is implicit and external to the 
Bitcoin blockchain. On the other hand, the Ethereum model explicitly keeps track of account values, and so this 
information needs to be kept as state information on chain. While it is possible to translate between UTxO-based and 
account-based chains \sjtnote{cite Bruno paper} the UTxO model is more fundamental, and requires less support from the 
implementation of the chain itself.

Contracts can be conceived of as acting in two ways. In a \emph{push} model, contracts are seen to be pro-active, and to 
make things happen. In the case of blockchain, this would include making payments or transactions happen. 
Alternatively, in the \emph{pull} model, contracts enable certain things to happen, such as transactions, but the 
transactions need to be effected by an external actor. The pull model makes fewer demands on the blockchain 
implementation than the push model, which requires the contract to be triggered automatically.

\subsection{Designing Marlowe}

In designing Marlowe, we have chosen a number of options; this description applies to Marlowe version 1.3.

\paragraph{Pull}

While earlier versions used a ``push'' model of operations, we have revised the language to give a ``pull'' semantics. 
This means, for example, that the language can \emph{enable} a payment to be made, but cannot make the payment happen. 
It is up to the interested party (or parties) to make the change happen. It is therefore possible for the payment not to 
happen, despite the fact that this is against the interest of the receiving party.

\paragraph{UTxO vs.\ accounts}

The language semantics is described in terms of \emph{commitments} of sums of cryptocurrency -- which we choose to 
abbreviate to ``cash'' or ``money'' -- and so is agnostic of the underlying implementation.  However, the current 
redemption mechanism in Plutus Core is not strong enough to support Marlowe, and we are actively investigating the best 
way to extend this mechanism to implement Marlowe in full on Plutus Core.

\paragraph{DSL or EDSL?}

Marlowe is described as an algebraic \haskellinline{data} type in Haskell, and as such can be seen as a simple EDSL. 
Note, however, that the implementation presented here is for the purposes of presenting the semantics and behaviour of 
the language in the abstract. The implementation on the Cardano platform could be stand-alone, but more likely it will 
be a DSL embedded in Plutus or Plutus Core. The details of this remain to be decided, as noted in the introduction.

@@

\begin{figure}[t]
\begin{centering}
\includegraphics[width=1.1\textwidth]{pix/screenshot1}
\par\end{centering}
\caption{\label{fig:full-screenshot-demo}Screenshot of Meadow tool
simulating the ``deposit incentive'' contract.}
\vspace*{-8mm}
\end{figure}

@@

@@

\subsection{Towards an implementation in Cardano}
\label{section:towards-implementation}

The contracts implemented in the Marlowe language have an implicit interface defined for what would correspond to the 
client side of a distributed application, as shown by the interactive interface.

We have designed Marlowe assuming that each contract would be compiled into
\begin{itemize} 
\item one or more smart contracts (potentially written in Plutus) and 
\item a client side program or interface that has a limited number of choices to be made at each point in time 
(including an interface for the observables). 
\end{itemize}
Of course, we probably do not want to give a user a programmatic interface but an actual application with user 
interface. So there is still some work to be done to create interfaces on top of the client side of the contracts.

Since the current version of Marlowe follows the `pull' model, a contract written in Marlowe could also be potentially
compiled to a smart-contract written in Ethereum. This could be done in two ways, it could be compiled so that an 
smart-contract represents a single instance of the Marlowe contract (with a fixed set of participants), or 
alternatively it would be possible to create a generic smart-contract with an extra ``instantiate'' operation that 
sets-up a new Marlowe contract instance with a given set of participants (represented by a set of public keys).

In principle, we can directly translate each operation in the model (i.e: cash commitment, payment claim, choice, or
redemption) into a separate transaction in the blockchain. In Ethereum model this would work in the same 
way that is simulated by Meadow: several operations can be issued in parallel and the blockchain subsystem will 
automatically apply them in some order and integrate them into the next block.

Another approach would be to use a UTXO model with continuations in which a UTXO represents a contract. In this model,
an operation would correspond to a transaction that spends a UTXO and creates (at least) a new UTXO that represents the 
contract after the operation.

Usually, in the UTXO model, it is necessary that a UTXO is in the blockchain in order for another transaction to be able 
to spend it. This could potentially limit the number of operations that can be applied in a single block to one, which 
would make infeasible the approach of using the same UTXO to represent several instances of the same contract (since it 
would not scale).

On the other hand, it would prevent non-determinism and, with it, it would remove the possibility of race-conditions. A 
change to the state of a contract before a transaction (and its corresponding operation) is accepted in the blockchain 
would invalidate the transaction and require the user to issue it again.

The Cardano platform has a notion of cost, and Marlowe contracts will potentially incur costs of two kinds:
\begin{itemize} 
\item The cost of a single transaction execution (a single Plutus validator).
\item The cost of the whole contract, which may consist of many transactions.
\end{itemize} 
Even in Ethereum this distinction is important because single contract execution is limited by gas (and is thus 
finite), whereas the life of the whole contract may be unbounded and consist of an indefinite number of transactions.  
So far we are assuming contracts defined in Marlowe to be finite in both aspects. They could be infinite if we were to 
define them co-inductively as recursive members of the DSL type.

If we have contracts that involve, for example, the oil price then we may need to convert a published price in USD into 
ADA, using a value for the prevailing USD/ADA exchange rate. This will require a computation: in this case, a 
multiplication. If Marlowe is to be stand-alone then we need to be extend it with arithmetic and other operations, but 
we expect Marlowe to be embedded in Plutus (Core) and so to be able to use the facilities of the host language here.

>>>>

We make the following contributions in this paper.
\begin{itemize}
\item
Designing a DSL for financial contracts on blockchains: Marlowe.
\item
Defining an executable, small-step semantics of Marlowe in Haskell.
\item 
Embedding Marlowe as a Haskell \haskellinline{data} type to extend its expressibility.
\item
Developing the Meadow tool that allows users to interact with and simulate the operation of embedded Marlowe contracts. 
\end{itemize}
Having established this model and its semantics we are able to do a number of other things, which we discuss in the paper. We can explore how the language will be implemented on an existing blockchain, such as Cardano, and our model has shown us that we will need to consider how the evolution of a contract interacts with the blockchain, with miners in the case of a `proof of work'-based chain, and with users participating in contract execution. We can also perform analyses of Marlowe contracts, based on its formal semantics. Finally, we can explore the design space for Marlowe, and explore potential alternative designs.

@@

TO UPDATE \ldots we begin in Section \ref{section:design} with a discussion of the design rationale 
for Marlowe, including an overview of the various approaches to scripting in blockchain as well as the different designs 
of the blockchains themselves. We conclude Section \ref{section:design} with a set of decisions about the blockchain 
environment for Marlowe.

@@

@@

\begin{figure}[t]
\begin{centering}
\includegraphics[width=1.1\textwidth]{pix/screenshot1}
\par\end{centering}
\caption{\label{fig:full-screenshot-demo}The Meadow tool
simulating the ``deposit incentive'' contract.}
\vspace*{-6mm}
\end{figure}

@@

\section{Implementation on blockchain}
\label{section:towards-implementation}

Marlowe is defined by its executable semantics, but to be deployed on blockchain it will have to be implemented on an existing distributed infrastructure.  We intend to deploy it on IOHK's Cardano infrastructure, but it can be implemented in other systems too, as we explain now.

\subsection{Abstraction}

Marlowe abstracts away from a number of concrete details of how blockchains operate. In particular, it is agnostic between UTxO-based systems, such as Bitcoin and Cardano SL, and account-based models such as Ethereum.

The simplest blockchain is Bitcoin. Transactions on Bitcoin are made by spending the (as yet) unspent outputs of 
previous transactions (`unspent transaction outputs' or UTxOs). The chain need not maintain any state, such as the 
current value owned by any particular participant: such \emph{account} information is implicit and external to the 
Bitcoin blockchain. On the other hand, the Ethereum model explicitly keeps track of account values, and so this 
information needs to be kept as state information on chain. While it is possible to translate between UTxO-based and 
account-based chains the UTxO model is more fundamental, and requires less support from the 
implementation of the chain itself.

Contracts can be conceived of as acting in two ways. In a \emph{push} model, contracts are seen to make things happen. In the case of blockchain, this would include making payments or transactions take plade. 
Alternatively, in a \emph{pull} model, contracts enable certain things, such as transactions,  to happen but the 
transactions need to be effected by an external actor. The pull model makes fewer demands on the blockchain implementation than the push model.

\subsection{Compilation}

We have designed Marlowe assuming that each contract would be compiled into
\begin{itemize} 
\item one or more on-chain programs and 
\item a client side program or interface that has a limited number of choices to be made at each point in time (including an interface for the observables). 
\end{itemize}
The contracts implemented in the Marlowe language have an implicit interface defined for what would correspond to the client side of a distributed application, as shown by the interactive interface. 
Of course, we probably do not want to give a user a programmatic interface but an actual application with user interface. So there is still some work to be done to create interfaces on top of the client side of the contracts.

A contract written in Marlowe could also be potentially
compiled to a smart-contract written in Ethereum. This could be done in two ways, it could be compiled so that an 
smart-contract represents a single instance of the Marlowe contract (with a fixed set of participants), or 
alternatively it would be possible to create a generic smart-contract with an extra ``instantiate'' operation that 
sets-up a new Marlowe contract instance with a given set of participants (represented by a set of public keys).

In principle, we can directly translate each operation in the model (i.e: cash commitment, payment claim, choice, or
redemption) into a separate transaction in the blockchain. In the Ethereum model this would work in the same 
way that is simulated by Meadow: several operations can be issued in parallel and the blockchain subsystem will 
automatically apply them in some order and integrate them into the next block.

Another approach would be to use a UTxO model with continuations in which a UTxO represents a contract. In this model,
an operation would correspond to a transaction that spends a UTxO and creates (at least) a new UTxO that represents the 
contract after the operation.

Usually, in the UTxO model, it is necessary that a UTxO is in the blockchain in order for another transaction to be able to spend it. This could potentially limit the number of operations that can be applied in a single block to one, which would make infeasible the approach of using the same UTxO to represent several instances of the same contract (since it would not scale).
On the other hand, it would prevent non-determinism and, with it, it would remove the possibility of race-conditions. A change to the state of a contract before a transaction (and its corresponding operation) is accepted in the blockchain 
would invalidate the transaction and require the user to issue it again.

The Cardano platform has a notion of cost, and Marlowe contracts will potentially incur costs of two kinds: The cost of a single transaction execution (e.g.\ a single Plutus validator), and the cost of the whole contract, which may consist of many transactions.

Even in Ethereum this distinction is important because single contract execution is limited by gas (and is thus 
finite), whereas the life of the whole contract may be unbounded and consist of an indefinite number of transactions.  
So far we are assuming contracts defined in Marlowe to be finite in both aspects. 
%They could be infinite if we were to 
%define them co-inductively as recursive members of the DSL type.

If we have contracts that involve, for example, the oil price then we may need to convert a published price in USD into 
ADA, using a value for the prevailing USD/ADA exchange rate. This will require a computation: in this case, a 
multiplication. If Marlowe is to be stand-alone then we need to be extend it with arithmetic and other operations, but 
we expect Marlowe to be embedded in a suitable language and so to be able to use the facilities of the host language here.

@@
